\documentclass[12pt]{article}
\usepackage{mgates-letter}
\definecolor{dark_blue} {rgb}{0., 0., 0.65}

\usepackage{textcomp}
\usepackage{mathrsfs}  % mathscr font
\usepackage{boxedminipage}
\usepackage{rotating}
\usepackage[inline]{enumitem}
%\usepackage{natbib}
\usepackage{xcolor}
\usepackage[colorlinks, filecolor=dark_blue, urlcolor=dark_blue, linkcolor=black, citecolor=black]{hyperref}

\newcommand{\todoinline}[1]{{\color{violet} #1}}

\begin{document}

\begin{titlepage}

	\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}
	\center
	
	\textsc{\Large Ph.D. Programme in Computer Science And Engineering}\\[0.5cm]
	
	\textsc{\Large XXXIX Cycle}\\[0.6cm]
	
	\hrule width \hsize \kern 1mm \hrule width \hsize height 2pt 
	\vspace{0.8cm}
	{ \large \bfseries Ph.D. Period Abroad Final Report for Marco Polo}\\[0.6cm]
	{ \large Transfer Learning and Coordination Techniques for Distributed Collective Intelligence }\\[0.6cm]

	\bfseries{December, 2025}


    \vspace{1.5cm}
    
    \noindent
    \begin{minipage}[t]{0.3\textwidth}
        \raggedright
        \textbf{Supervisors:}\\[0.5cm]
        Prof. Mirko Viroli\\
        Prof. Danilo Pianini\\
        Prof. Matteo Ferrara
    \end{minipage}%
	\hfill
    \begin{minipage}[t]{0.3\textwidth}
        \centering
        \textbf{Abroad Supervisor:}\\[0.5cm]
        Ivana Dusparic 
    \end{minipage}
    \hfill
    \begin{minipage}[t]{0.3\textwidth}
        \raggedleft
        \textbf{PhD Candidate:}\\[0.5cm]
        Davide Domini
    \end{minipage} \\[0.6cm]

	\hrule width \hsize height 2pt \kern 1mm \hrule width \hsize height 1pt
	\vspace{0.4cm}

\end{titlepage}

\section{The stay}
During my research stay abroad, 
 carried out at Trinity College Dublin (TCD) from September to December 2025, 
 I collaborated with the Distributed Systems Group under the supervision of Prof. Ivana Dusparic.
%
The main objective of the stay was to advance joint research activities in the area 
 of transfer learning for Multi-Agent Reinforcement Learning (MARL), 
 with a particular focus on distributed and decentralized settings.

This period at TCD represented a valuable opportunity to align my PhD research with the expertise and 
 ongoing work of the host group. 
% 
The collaboration was fully consistent with the objectives outlined in my period abroad proposal and 
 enabled the exploration of new methodological perspectives, 
 as well as the establishment of promising research synergies.

\section{Research context and objectives}
The research activity conducted during the stay was situated within the broader context 
 of multi-agent systems and distributed learning, 
 focusing on the role of transfer learning as a mechanism to improve scalability, stability, and 
 learning efficiency in MARL scenarios characterized by heterogeneity and dynamic environments.

In particular, 
 the goal was to investigate how knowledge transfer mechanisms can be effectively integrated 
 into decentralized multi-agent learning architectures, 
 enabling agents to reuse experience acquired by others while reducing redundant learning. 
% 
The interaction with the Distributed Systems Group also allowed these topics to be analyzed 
 from a distributed systems perspective, highlighting connections between learning, coordination, 
 and communication.

\section{Main activities}
During my stay at Trinity College Dublin, I carried out the following main activities:
\begin{itemize}
	\item regular research meetings with Prof. Ivana Dusparic and members of the Distributed Systems Group,
	 aimed at discussing ideas, methodologies, and potential joint research directions;
	\item in-depth study and analysis of the relevant literature on transfer learning and 
	 multi-agent reinforcement learning, with particular emphasis on coordination and generalization 
	 challenges;
	\item initiation of preliminary experimental work exploring the intersection between 
	 my PhD research topics and the research interests of the host group; 
	 these experiments are currently still under development;
	\item active participation in the academic life of the group through attendance at seminars and 
	 research meetings.
\end{itemize}

\section{Seminars and dissemination activities}
In addition to research activities,
 I was involved in several seminar-related initiatives. 
%
In particular, 
 I delivered an internal seminar to the research group at TCD, 
 presenting my PhD research topics and ongoing work. 
% 
This seminar fostered constructive discussion and provided valuable feedback from PhD students and 
 researchers within the group.

I also attended multiple seminars organized by the department, 
 including a presentation by one of the group's PhD students on a paper recently published at ECAI, 
 focused on Multi-Objective Reinforcement Learning (MORL). 
% 
These activities contributed to broadening my understanding of related research areas and 
 strengthening scientific exchange within the host institution.

\section{Ongoing and future work}
Throughout the stay, 
 several points of connection emerged between my PhD research and the work of 
 the Distributed Systems Group. 
% 
In particular, 
 we began exploring how transfer learning techniques could be combined with coordination mechanisms 
 in distributed multi-agent systems.

These discussions led to the development of initial experimental setups that are currently still ongoing. 
%
Although no finalized results are available yet, 
 the work carried out during the stay has laid the foundations for a continued collaboration, 
 which is expected to result in one or more joint publications in the coming months.


%\bibliographystyle{unsrt}
%\bibliography{bibliography}

\end{document}
